A Systematic Literature Review on Extensions to Role-Based Access Control

### 2014-02-09 paper/sections/intro.tex
Updating data display tables.
update intro
Add space before citation []

administration[? ] -> administration [? ]
the 1990s[? ], again in 2002[? ] -> the 1990s [? ], again in 2002 [? ]
and later confirmed in 2010[? ]. -> and later confirmed in 2010 [? ].

Software systems use access control mechanisms to determine which subjects can access which resources.
Role-Based Access Control (RBAC) is a widely used access control mechanism designed for maintaining and managing an organization's  access control based on assigning permissions to roles, and roles to users, instead of assigning the permissions directly to individual users.
RBAC is used for securing various applications including web services, database applications, and healthcare applications. 
In RBAC, roles represent a set of permissions needed to perform a particular job function within an organization.  
Multiple users who are involved in that specific job function within the organization can then be assigned to a single role to inherit the required access. 
The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows and the number of permissions and users scales upward.
As permissions can be managed by role instead of by user, RBAC has been shown to significantly reduce complexity of security administration\cite{impact2002}.
For example, if a user requires access to resources associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead of assigning a set of individual permissions.

The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in 2000 \cite{sandhu2000nist}.
NIST requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu et al.~\cite{sandhu1996role}.  
In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute\footnote{http://www.ansi.org/} (ANSI) and the InterNational Committee for Information Technology Standards\footnote{http://www.incits.org/} (INCITS).
The development of a standard was inspired by an economic impact study done during the 1990s\cite{impact1996}, again in 2002\cite{impact2002} and later confirmed in 2010\cite{o20102010}. 
The study showed the cost savings of RBAC implementation and maintenance. 
Prior to the development of the RBAC standard, vendors proposed and implemented their own RBAC definition without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles). 
The RBAC standard includes the RBAC Reference Model which serves as a basis for defining the scope and functional specifications of RBAC features.

@@

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item A starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
the current state of the RBAC extension models. 
\item A basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs.
\item A summary of current evaluation methods used in research on RBAC extension models.
\item A starting direction for a new standard based on conclusions drawn from the RBAC extension models.
\end{itemize}

>>>>

Software systems use access control mechanisms to determine which subjects can access which resources.
Role-Based Access Control (RBAC) is a widely used access control mechanism designed for maintaining and managing an organization's  access control based on assigning permissions to roles, and roles to users, instead of assigning the permissions directly to individual users.
RBAC is used for securing various applications including web services, database applications, and healthcare applications. 
In RBAC, roles represent a set of permissions needed to perform a particular job function within an organization.  
Multiple users who are involved in that specific job function within the organization can then be assigned to a single role to inherit the required access. 
The ability to logically group users into roles associated with permissions becomes paramount for managing access control as an organization grows and the number of permissions and users scales upward.
As permissions can be managed by role instead of by user, RBAC has been shown to significantly reduce complexity of security administration \cite{impact2002}.
For example, if a user requires access to resources associated with a manager role within a given organization, security policy administrators need only associate the user with the manager role instead of assigning a set of individual permissions.

The use of RBAC has become popular since the National Institute for Standards and Technology (NIST) first proposed the RBAC standard in 2000 \cite{sandhu2000nist}.
NIST requested that a unified standard be created by combining the Ferraiolo and Kuhn model \cite{ferraiolokuhn} with the framework proposed by Sandhu et al.~\cite{sandhu1996role}.  
In 2004, this standard was adopted as ANSI/INCITS 359-2004 approved by American National Standards Institute\footnote{http://www.ansi.org/} (ANSI) and the InterNational Committee for Information Technology Standards\footnote{http://www.incits.org/} (INCITS).
The development of a standard was inspired by an economic impact study done during the 1990s \cite{impact1996}, again in 2002 \cite{impact2002} and later confirmed in 2010 \cite{o20102010}. 
The study showed the cost savings of RBAC implementation and maintenance. 
Prior to the development of the RBAC standard, vendors proposed and implemented their own RBAC definition without general agreement on a unified definition of RBAC or RBAC features (e.g., inheritance relationships among roles). 
The RBAC standard includes the RBAC Reference Model which serves as a basis for defining the scope and functional specifications of RBAC features.

@@

\begin{itemize}
\setlength{\itemsep}{0.25pt}
\item A starting place for researchers looking to tackle new problems in the realm of authorization and as a reference guide for discovering
the current state of the RBAC extension models. 
\item A basis for comparison and look up for what extension model to use for developers looking to find a model to fit their access control needs.
\item A summary of current evaluation methods used in research on RBAC extension models.
\item A starting point for a new standard based on conclusions drawn from the RBAC extension models.
\end{itemize}
